%!TEX root = draft.tex
\begin{abstract}
  %We address the problem of specifying and verifying distributed 
  
  Geo-distributed systems often replicate data at multiple locations to achieve availability and performance, despite network partitions. 
%  
  These systems must accept updates at any replica, and propagate these updates asynchronously to the other replicas.
%  
  Convergent Replicated Data Types (CRDTs) provide a principled approach to the problem of ensuring that replicas are eventually consistent despite the asynchronous delivery of updates.

  We address the problem of specifying and verifying CRDTs, introducing a new correctness criterion called Replication-Aware Linearizability. This criterion is inspired by linearizability, the de-facto correctness criterion for (shared-memory) concurrent data structures.
%
%  In this paper we consider the problem of the specification and
%  verification of Commutative Replicated Data Types (CRDTs).
%  %
%  We provide a new specification and correctness criterion for CRDTs
%  akin to Linearizability as defined by~\citet{HerlihyW90}.
  %
  We argue that this criterion is both simple to understand, and it
  fits most known implementations of CRDTs.
  %
  We provide a proof methodology for showing that a CRDT satisfies replication-aware linearizability
  which we apply on a wide range of implementations.
%  Then we show how to prove that CRDT implementations can be
%  formally proved to be correct w.r.t. their specification.
%  %
%  In particular, we do so for many well known implementations taken
%  from~\citet{ShapiroPBZ11}.
  %
  Finally, we show that our criterion can be leveraged
  to reason modularly about the composition of CRDTs.
\end{abstract}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft"
%%% End:
