%!TEX root = draft.tex
%\newcommand{\seqPQ}{\mathsf{SeqPQ}}

\section{CRDT Implementation Correctness}
\label{sec:CRDT implementation semantics and correctness}

\textblue{
I think this section should be about proving linearizability. Here we introduce:
\begin{itemize}
\item a semantics for CRDTs
\item a general methodology based on simulation relation (explain that it can also be seen as invariant checking)
\item the two proof strategies for building linearizations, as particular instances of that simulation relation
\end{itemize}}


In this section, we propose the semantics of CRDT implementations. Based on our definition, we can extract histories from an execution. Then we define the correctness of an CRDT implementation.



%\subsection{Semantics of Single Object}
%\label{subsec:semantics of single object}

CRDT implementations assume the following two guarantees of downstream:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] The downstream of each operation is applied at most once for each replica.
\item[-] Assume the downstream of operation $\aop_1$ is applied in replica of $\aop_2$ before $\aop_2$ happens. Then, for each replica $\arep$, the downstream of $\aop_2$ can be applied only if the downstream of $\aop_1$ has already been applied.
\end{itemize}

Given a CRDT object $\aobj$, its semantics is defined as a transition system $\llbracket \aobj \rrbracket = (\globalstates,\aglobalstate_0,\labels,\rightarrow)$ as in \figurename~\ref{fig:crdt-opsem}.



%\gpfatal[nomargin, inline]{remove if we decide to keep the other semantics}

\begin{figure}[t]
  \centering

\begin{itemize}
\item $ \Theta : \states \times \methods \times \datadomain \times \reps \Rightarrow \datadomain \times \Delta \times \tss \ni \atsource$ \hspace{\fill} AtSource 
\item $ \Delta : \states \rightarrow \states \ni \effector$ \hspace{\fill} Effector or Downstream
%\item $\semop[]{} : \labels \times \states \rightarrow (\states \rightarrow \states)$ \hspace{\fill} Operational Semantics of a Single Operation (Label)
\item $\localstates : \powerset{\labels} \times \states \ni (\alabelset, \astate)$ \hspace{\fill} Local Configurations
\item $\globalstates : (\reps \rightarrow \localstates) \times \powerset{\labels \times \labels} \times (\labels \rightarrow \Delta) \ni (\gstates, \avisord, \downstreams )$ \hspace{\fill} Global Configurations
\end{itemize}


\[
  \inferrule[\text{\sc Operation}]
  {\gstates(r) = (\alabelset, \astate) \\ \theta(\sigma,\amethod,\argv,\arep) = (\retv,\effector,\ats) \\  \effector(\astate) = \astate' \\ \alabel = \alabellongind{\argv}{\retv}{(i,\ats)} \\ \mathit{unique}(i) }
  {(\gstates, \avisord, \downstreams) \xrightarrow{\alabel} (\gstates[\arep \leftarrow (\alabelset \cup \{\alabel\}, \astate')],
    \avisord \cup (\alabelset \times \{\alabel\}), \downstreams[\alabel \rightarrow \effector])}
\]


\[
  \inferrule[\text{\sc DownStream}]
  {\gstates(r) = (\alabelset, \astate) \\ \alabel \in \mathsf{min}_{\avisord}(\dom{\avisord} / \alabelset) \\
    \downstreams(\alabel)(\astate) = \astate'}
  {(\gstates, \avisord, \downstreams) \xrightarrow{} (\gstates[\arep \leftarrow (\alabelset \cup \{\alabel\}, \astate')], \avisord, \downstreams)}
\]

  \caption{Operational Semantics of CRDTs}
  \label{fig:crdt-opsem}
\end{figure}

A global configuration $(\gstates, \avisord, \downstreams)$ is a snapshot of distributed system. $\gstates \in \reps \rightarrow \localstates$ stores the local configuration of each replica. A local configuration $(\alabelset, \astate)$ contains a local state $\astate$ of $\aobj$ and a set $\alabelset$ of labels: $\alabel \in \alabelset$, if the downstream of $\alabel$ has already been applied in this replica. $\avisord \in \powerset{\labels \times \labels}$ stores the visibility among operations, and $\downstreams$ stores the downstream of operations. $\aglobalstate_0 = (\gstates_0, \emptyset, \emptyset) \in \globalstates$ is the initial global configuration, where $\gstates_0$ maps each replica $\arep$ into $(\emptyset, \astate_0)$ and $\astate_0$ is the initial local state. 

$\rightarrow \in \globalstates \times \labels \times \globalstates$ is the transition relation. The first rule in \figurename~\ref{fig:crdt-opsem} describes replica $\arep$ performing an operation and generate a new downstream. $\mathit{unique}(i)$ ensures that $i$ is a unique identifier, and then $\alabel$ is a unique operation label. A new downstream $\effector$ is generated and applied to replica $\arep$. The set $\alabelset$ and $\avisord$ is updated according to $\alabel$. Note that here if an operation generates a new timestamp, then it is recorded in the label of operation $\alabel$. The second rule describes the downstream of $\alabel$ being applied in a replica that is not the replica launches $\alabel$. With $\alabel \in \mathsf{min}_{\avisord}(\dom{\avisord} / \alabelset)$, we always choose a operation $\alabel$ that is minimal w.r.t $\avisord$ among operations whose downstream not applied to replica $\arep$ yet. This satisfies the guarantees of downstream.


A sequence $\aexec$ of operation labels is an execution of $\llbracket \aobj \rrbracket$, if there exists a global configuration $\aglobalstate = (\gstates, \avisord, \downstreams)$, such that $\aglobalstate_0 {\xrightarrow{ l }} \aglobalstate$. The history of $\aexec$ is a tuple $(\alabelset, \avisord)$, where $\alabelset$ is the set of labels of $l$. Let $\mathit{history}(\llbracket \aobj \rrbracket)$ be the set of histories of all executions of $\llbracket \aobj \rrbracket$. Then, $\aobj$ is \crdtlinearizable{}, if each of its history is, as stated by the following:

\begin{definition}[Correctness of a CRDT Object]
\label{definition:correctness of a CRDT object}
A CRDT object $\aobj$ is \crdtlinearizable{} w.r.t a sequential specification \Spec{}, if each history of $\mathit{history}(\llbracket \aobj \rrbracket)$ is \crdtlinearizable{} w.r.t \Spec{}.
\end{definition}

















\forget{
In this section, we propose the semantics of %a state-based CRDT or a operation-based CRDT.
CRDT implementations. Then we shows how to extract histories from execution, and the correctness of CRDT implementations.



\subsection{Semantics of a Single Object}
\label{subsec:semantics of a single object}

CRDT implementations assume the following two guarantees of message delivery:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] No duplicated delivery: Each message is delivered to a replica at most once.
\item[-] Causal delivery: Given two update operations $o_1$ and $o_2$, and assume $o_1$ is visible to $o_2$. Or we can say, $o_1$ is related to $o_2$ by several times of replica order and message delivery. Then, for each replica, once it receives $o_2$, it must be the case that $o_1$ has been previously received.
\end{itemize}

Given a CRDT object $\mathit{obj}$ with $\Sigma$ be the set of local states, its semantics is defined as $\llbracket \mathit{obj} \rrbracket_{\mathit{op}} = (\mathit{Config},\mathit{config}_0,\Sigma',\rightarrow)$ as in \figurename~\ref{fig:the semantics of a operation-based CRDT object}.


\gpfatal[nomargin, inline]{remove if we decide to keep the other semantics}

\begin{figure}[t]
  \centering

\begin{itemize}
\item $ \Delta : \states \rightarrow \states \ni \effector$ \hspace{\fill} Effector or Downstream
\item $\semop[]{} : \labels \times \states \rightarrow (\states \rightarrow \states)$ \hspace{\fill} Operational Semantics of a Single Operation (Label)
\item $\localstates : \powerset{\labels} \times \states \ni (\alabelset, \astate)$ \hspace{\fill} Local States
\item $\globalstates : (\reps \rightarrow \localstates) \times \powerset{\labels \times \labels} \times (\labels \rightarrow \Delta) \ni (\gstates, \avisord, \downstreams )$ \hspace{\fill} Global States
\end{itemize}


\[
  \inferrule[\text{\sc Operation}]
  {\gstates(r) = (\alabelset, \astate) \\ \semop[\alabel]{\astate} = \effector \\ \effector(\astate) = \astate'}
  {(\gstates, \avisord, \downstreams) \xrightarrow{\alabel} (\gstates[\arep \leftarrow (\alabelset \cup \{\alabel\}, \astate', \downstreams)],
    \avisord \cup (\alabelset \times \{\alabel\}), \downstreams[\alabel \rightarrow \effector])}
\]

\[
  \inferrule[\text{\sc DownStream}]
  {\gstates(r) = (\alabelset, \astate) \\ \alabel \in \mathsf{min}_{\avisord}(\dom{\avisord} / \alabelset) \\
    \downstreams(\alabel)(\astate) = \astate'}
  {(\gstates, \avisord, \downstreams) \xrightarrow{} (\gstates[\arep \leftarrow (\alabelset \cup \{\alabel\}, \astate')], \avisord, \downstreams)}
\]

  \caption{Operational Semantics of CRDTs}
  \label{fig:crdt-opsem}
\end{figure}


\begin{figure}[ht]
$\mathit{RState} = \mathbb{R} \rightarrow \Sigma$

$\mathit{TState} = \mathbb{MID} \times \mathbb{MSG} \times \mathbb{R}$

$\mathit{MsgHB} \subseteq \mathbb{MID} \times \mathbb{MID}$

$\mathit{MsgDel} \subseteq \mathbb{MID} \times \mathbb{R}$

$\mathit{Config} = \mathit{RState} \times \mathit{TState} \times \mathit{MsgHB} \times \mathit{MsgDel}$, $\mathit{config}_0 \in \mathit{Config}$.

$\Sigma' = \mathit{do}(\mathbb{M} \times \mathbb{D} \times \mathbb{D} \times \mathbb{R} \times \mathbb{MID}) \cup \mathit{do}(\mathbb{M} \times \mathbb{D} \times \mathbb{D} \times \mathbb{R}) \cup \mathit{receive}(\mathbb{MID} \times \mathbb{R})$

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[] $\begin{array}{l c}
   \bigfrac{
   \begin{array}{c}
     R(r) = \sigma, r.\mathit{do}(\sigma,m,a) = (\sigma',b,\mathit{msg}), \mathit{msg} \neq \emptyset, \mathit{unique}(\mathit{mid}), \\
     S_1 = \{ (\mathit{mid}',\mathit{mid}) \vert (\mathit{mid'},r) \in \mathit{MsgDel} \}, S_2 = \{ (\mathit{mid}',\mathit{mid}) \vert \mathit{mid}' \in T, \mathit{mid}' \ is \ of \ replica \ r \}
   \end{array}}
     {(R,T,\mathit{msgHB},\mathit{MsgDel}) {\xrightarrow{\mathit{do}(m,a,b,r,\mathit{mid})}} (R[r:\sigma'], T \cup \{ (\mathit{mid},\mathit{msg},r) \}, (\mathit{MsgHB} \cup S_1 \cup S_2)^*,\mathit{MsgDel})}
\end{array}$

\item[] $\begin{array}{l c}
   \bigfrac{
   \begin{array}{c}
     R(r) = \sigma, r.\mathit{do}(\sigma,m,a) = (\sigma',b,\emptyset)
   \end{array}}
     {(R,T,\mathit{msgHB},\mathit{MsgDel}) {\xrightarrow{\mathit{do}(m,a,b,r)}} (R[r:\sigma'], T \cup \{ (\mathit{mid},\mathit{msg},r) \}, \mathit{MsgHB},\mathit{MsgDel})}
\end{array}$

\item[-] $\begin{array}{l c}
   \bigfrac{
   \begin{array}{c}
      R(r) = \sigma, r.\mathit{receive}(\sigma,\mathit{msg}) = \sigma', (\mathit{mid},\mathit{msg},r') \in T, r \neq r', \\
      (\mathit{mid},r) \notin \mathit{MsgDel}, \mathit{mid} \ is \ minimal \ w.r.t \ \mathit{MsgHB} \ among \ \{ \mathit{mid}' \vert (\mathit{mid}',r) \notin \mathit{MsgDel} \}
   \end{array}}
     {(R,T,\mathit{msgHB},\mathit{MsgDel}) {\xrightarrow{\mathit{receive}(\mathit{mid},r)}} (R,T,\mathit{MsgHB},\mathit{MsgDel} \cup \{ (\mathit{mid},r) \} )}
\end{array}$
\end{itemize}
\caption{The definition of semantics of $\llbracket \mathit{obj} \rrbracket_{\mathit{op}}$}
\label{fig:the semantics of a operation-based CRDT object}
\end{figure}

A configuration $(R,T,\mathit{MsgHB},\mathit{MsgDel})$ is a snapshot of distributed system. $R$ gives the local state of each replica, and $T$ gives the set of messages that has been generated. Let $\mathbb{MID}$ be the set of message identifiers of message content. A message is a tuple $(\mathit{mid},\mathit{msg},r)$, where $\mathit{mid} \in \mathbb{MID}$ is the identifier, $\mathit{msg} \in \mathbb{MSG}$ is the message content, and $r$ is the original replica of message. $\mathit{MsgHB}$ is used to record the happen-before relation between messages: two messages $(\mathit{mid}_1,\mathit{mid}_2) \in \mathit{MsgHB}$ represents that the operation of $\mathit{mid}_1$ happens before the operation of $\mathit{mid}_2$. $\mathit{MsgDel}$ is used to record the message delivery relation between messages: $(\mathit{mid},r) \in \mathit{MsgDel}$ represents that message $\mathit{mid}_1$ has already been delivered to replica $r$. $\mathit{MsgHB}$ and $\mathit{MsgDel}$ are used to ensure message delivery requirements. $\mathit{config}_0$ is the initial configuration, which maps each replica into the initial local state, has no message, and with a empty happen-before relation and a empty message delivery relation.


Each element of $\Sigma'$ is called an action. $\rightarrow \in \mathit{Config} \times \Sigma' \times \mathit{Config}$ is the transition relation and describes a single step of distributed systems. The first rule in \figurename~\ref{fig:the semantics of a operation-based CRDT object} describes replica $r$ performs a update operation $m(a) \Rightarrow b$ and generates a message with message content $\mathit{msg}$. Here $\mathit{unique}$ is a function that ensures $\mathit{mid}$ be a fresh message identifier. We insert message identifier $\mathit{mid}$ into the happen-before relation and keeps the happen-before relation be transitive. The second rule describes replica $r$ performs a query operation $m(a) \Rightarrow b$ and thus does not generate message. Since no message is generated, the $\mathit{MsgHB}$ and $\mathit{MsgDel}$ tuples remain the same. The third rule describes delivery of a message to a replica $r$ other than its origin replica $r'$. By $(\mathit{mid},r) \notin \mathit{MsgDel}$, we ensure that $\mathit{mid}$ has not been previously delivered to replica $r$, and thus, each message be delivered to each replica at most once. By $\mathit{mid}$ being minimal w.r.t $\mathit{MsgHB}$ among $\{ \mathit{mid}' \vert (\mathit{mid}',r) \notin \mathit{MsgDel} \}$, we always choose a minimal element w.r.t $\mathit{MsgHB}$ among operations not been delivered to a replica, and thus, ensures causal-delivery.



A sequence $l$ of actions is an execution of $\llbracket \mathit{obj} \rrbracket_{\mathit{op}} = (\mathit{Config},\mathit{config}_0,\Sigma',\rightarrow)$, if there exists $(R,T,\mathit{MsgHB},\mathit{MsgDel}) \in \mathit{Config}$, such that $\mathit{config}_0 {\xrightarrow{ l }} (R,T,\mathit{MsgHB},\mathit{MsgDel})$. The semantics of $\mathit{obj}$ is defined as the set of executions of $\llbracket \mathit{obj} \rrbracket_{\mathit{op}}$.




\subsection{Correctness of a Single Object}
\label{subsec:correctness of a single object}

Given an execution $l = \alpha_1 \cdot \ldots \cdot \alpha_k$ of $\llbracket \mathit{obj} \rrbracket_{\mathit{op}}$ of CRDT object $\mathit{obj}$, we can associate unique operation identifiers with each action of $l$, and then obtain a corresponding history $\mathit{history}(l) = (\mathit{Op},\mathit{ro},\mathit{vis})$, such that

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Each operation in $\mathit{Op}$ is a tuple $(m(a) \Rightarrow b,i,\mathit{obj})$, such that $i$ is the operation identifier of a $\mathit{do}(m,a,b,r)$ action of $l$.

\item[-] $(o_1,o_2) \in \mathit{ro}$, if they are of same replica, and the index of $o_1$ in $h$ is before that of $o_2$.

\item[-] $(o_1,o_2) \in \mathit{vis}$, if the action of $o_1$ is $\mathit{send}(\mathit{mid},\_)$, and either the action of $o_2$ is $\mathit{receive}(\mathit{mid},\_)$, or there exists $o_3$, such that the action of $o_3$ is $\mathit{receive}(\mathit{mid},\_)$, and $(o_3,o_2) \in \mathit{ro}$.
\end{itemize}

Let $\mathit{history}(\llbracket \mathit{obj} \rrbracket_{\mathit{op}})$ be the set of histories of all executions of $\llbracket \mathit{obj} \rrbracket_{\mathit{op}}$. Then, a object is distributed linearizable, if each of its history is, as stated by the following:

\begin{definition}[Correctness of a CRDT Object]
\label{definition:correctness of a CRDT object}
A CRDT object $\mathit{obj}$ is distributed linearizable w.r.t a sequential specification $\mathit{spec}$, if each history of $\mathit{history}(\llbracket \mathit{obj} \rrbracket_{\mathit{op}})$ is distributed linearizable w.r.t $\mathit{spec}$.
\end{definition}
}





















\forget
{
\subsection{Semantics}
\label{subsec:semantics}

Given a set $\mathit{Obj}$ of objects, we define its semantics as a set of executions generated from an LTS $\llbracket \mathit{Obj} \rrbracket = (\mathit{Config},\mathit{config}_0,\Sigma',\rightarrow)$ as in \figurename~\ref{fig:the semantics of multiple objects}.

\begin{figure}[ht]
$\mathit{RState} = \cup_{x \in \mathit{Obj}} (\mathbb{R} \rightarrow x.\Sigma)$

$\mathit{TState} = \mathbb{MID} \times \mathbb{MSG} \times \mathit{Obj} \times \mathbb{R}$

$\mathit{Config} = \mathit{RState} \times \mathit{TState}$, $\mathit{config}_0 \in \mathit{Config}$.

$\Sigma' = \mathit{do}(\mathit{Obj} \times \mathbb{M} \times \mathbb{D} \times \mathbb{D} \mathbb{MID}) \cup \mathit{receive}(\mathit{Obj} \times \mathbb{MID} \times \mathbb{R})$

\[
\begin{array}{l c}
\bigfrac{ R(x,r) = \sigma, x(r).\mathit{do}(\sigma,m,a) = (\sigma',b,\mathit{msg}), \mathit{msg} \neq \emptyset, \mathit{unique}(\mathit{mid}) }
{ (R,T) {\xrightarrow{\mathit{do}(x,m,a,b,r,\mathit{mid})}} (R[(x,r):\sigma'],T \cup \{ (\mathit{mid},\mathit{msg},x,r) \}) }
\end{array}
\]

\[
\begin{array}{l c}
\bigfrac{ R(x,r) = \sigma, x(r).\mathit{do}(\sigma,m,a) = (\sigma',b,\mathit{msg}), \mathit{msg} = \emptyset }
{ (R,T) {\xrightarrow{\mathit{do}(x,m,a,b,r)}} (R[(x,r):\sigma'],T ) }
\end{array}
\]

\[
\begin{array}{l c}
\bigfrac{ R(x,r) = \sigma, x(r).\mathit{receive}(\sigma,\mathit{msg}) = \sigma',(\mathit{mid},\mathit{msg},x,r') \in T, r \neq r'}
{ (R,T) {\xrightarrow{\mathit{receive}(x,\mathit{mid},r)}} (R[(x,r):\sigma'],T) }
\end{array}
\]
\caption{The definition of semantics of $\llbracket \mathit{Obj} \rrbracket$}
\label{fig:the semantics of multiple objects}
\end{figure}

A configuration $(R,T)$ is a snapshot of distributed system and contains two parts: $R$ gives the local state of each object at each replica, and $T$ gives the set of messages that has been generated. Let $\mathbb{MID}$ be the set of message identifiers of message content. A message is a tuple $(\mathit{mid},\mathit{msg},x,r)$, where $\mathit{msgId} \in \mathbb{MID}$ is the identifier, $\mathit{msg} \in \mathbb{MSG}$ is the message content, $x$ is the object this message pertains to, and $r$ is the original replica of message. $\mathit{config}_0$ is the initial configuration, which maps each object at each replica into its initial local state, and has no message inside.

Each element of $\Sigma'$ is called an action. $\rightarrow \in \mathit{Config} \times \Sigma' \times \mathit{Config}$ is the transition relation and describe a single step of distributed systems. The first rule in \figurename~\ref{fig:the semantics of multiple objects} describes replica $r$ performs a update operation $m(a) \Rightarrow b$ of object $x$ and generate a message with message content $\mathit{msg}$. Here $\mathit{unique}$ is a function that ensures $\mathit{mid}$ is a fresh message identifier. The first rule  describes replica $r$ performs a query operation $m(a) \Rightarrow b$ of object $x$ and thus does not generate message. The third rule describes delivery of a message of object $x$ to a replica $r$ other than its origin replica $r'$.

A sequence $l$ of actions is an execution of $\llbracket \mathit{Obj} \rrbracket = (\mathit{Config},\mathit{config}_0,\Sigma',\rightarrow)$, if there exists $(R,T) \in \mathit{Config}$, such that $\mathit{config}_0 {\xrightarrow{ l }} (R,T)$. The semantics of $\mathit{Obj}$ is defined as the set of executions of $\llbracket \mathit{Obj} \rrbracket$. Given an execution, when the context is clear, we can associate a unique operation identifier to each action. Or we can say, it is safe to assume each action is either $\mathit{do}(i,x,m,a,b,r,\mathit{mid})$ or $\mathit{receive}(i,x,\mathit{mid},r)$, where $i \mathbb{OID}$ is a unique operation identifier.





\subsection{Correctness of a CRDT Implementation}
\label{subsec:correctness of a CRDT implementation}

To state the correctness of a single CRDT implementation, we consider the semantics of a single CRDT object.



Note that the transition relation of $\llbracket \mathit{Obj} \rrbracket$ does not make any assumption about message delivery: Messages can be delivered in any order; a message can be delivery to a replica multiple times; and a message can be never delivered to a replica. However, although the state-based CRDT do not assume any guarantee about message delivery, the operation-based CRDT assume the following two guarantees about message delivery:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Each message is delivered to a replica at most once.
\item[-] Causal delivery: Given two update operations $o_1$ and $o_2$ of a same object, and assume $o_1$ is visible to $o_2$. Or we can say, $o_1$ is related to $o_2$ by several times of replica order and message delivery (all intermediate action is of the same object of $o_1$ and $o_2$). Then, for each replica, once it receives $o_2$, it must be the case that $o_1$ has been previously received.
\end{itemize}

Given an execution $l = \alpha_1 \cdot \ldots \cdot \alpha_k$ of $\llbracket \mathit{Obj} \rrbracket$,
}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft"
%%% End:
