\documentclass{llncs}
\usepackage[english]{babel}

\usepackage{hyperref}

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{color}
\usepackage{amsfonts,amssymb}
\usepackage{mathabx}
\usepackage{verbatim}
\usepackage{times}
\usepackage[linesnumbered,ruled,procnumbered,noend]{algorithm2e}
\usepackage{stmaryrd}

%added by Wang Chao%
%\usepackage{mathrsfs}
%\usepackage{extarrows}
\usepackage{thmtools}
\usepackage{thm-restate}

%add for TIkz
\usepackage[version=0.96]{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}
\usepackage{xcolor}
\usepackage{paralist}

\usepackage[margin,draft]{fixme}
\fxsetup{theme=color,mode=multiuser}
\FXRegisterAuthor{gp}{agp}{GP}
\FXRegisterAuthor{cw}{acw}{CW}

\newcommand{\gpn}[2][]{\gpnote*{#1}{#2}}
\newcommand{\gp}[1]{\gpnote{#1}}


\DeclareSymbolFont{largesymbolsA}{U}{txexa}{m}{n}
\SetSymbolFont{largesymbolsA}{bold}{U}{txexa}{bx}{n}
\DeclareFontSubstitution{U}{txexa}{m}{n}
\DeclareMathSymbol{\bigsqcupplus}{\mathop}{largesymbolsA}{"02}

\title{CRDT Linearizability}

\author{}
\institute{}

\author {Constantin Enea \and Gustavo Petri \and Chao Wang}
\institute{Institut de Recherche en Informatique Fondamentale, \\Univ. Paris Diderot (Paris 7)}

\input{macros}

\begin{document}

\maketitle

\input{Abstract}

\input{Introduction}

\input{CRDT-Implemtation}

\input{LinDef}

\input{CRDT-Semantics}

\input{Linearization-Proof}

%\section{CRDT-Linearizable Implementations}
%\label{sec:crdt-lin-imp}

%\gpn{Proofs of CRDT-Linearizable implementations}

%\begin{itemize}
%\item Counters
%  \begin{itemize}
%  \item Grow-Only Counter
%  \item PN-Counter
%  \end{itemize}
%\item Registers:
%  \begin{itemize}
%  \item LWW
%  \item MVR
%  \end{itemize}
%\item Sets:
%  \begin{itemize}
%  \item LAW (OR-Set)
%  \item Grow-Only Set
%  \end{itemize}
%\item Text-Editing (Graphs/Lists)
%  \begin{itemize}
%  \item RGA (addRight)
%  \item WOOT
%  \end{itemize}
%\end{itemize}

\section{Meta-properties of CRDT-Linearizability}
\label{sec:meta-prop-lincrdt}

\subsection{Compositionality}
\label{sec:compositionality}

\gpn{Here we explain the compositionality result a la Herlihy Wing}

\begin{itemize}
\item What are the conditions for compositionality (deterministic
  specs?, conflict?)
  \begin{itemize}
  \item Can we share the same timestamp for different data structures.
  \item Coherent conflict resolution.
  \end{itemize}
\item Should we change the implementations to change the
  conflict-resolution.
\end{itemize}

\subsection{Abstraction}
\label{sec:abstraction}

\gpn{Here we explain refinement}
\begin{itemize}
\item What is the power of the client? Messages, other calls to CRDTs?
\item Observational refinement?
\item Completeness?
\end{itemize}

\bibliographystyle{plainurl}
\bibliography{draft}

\begin{thebibliography}{50}
\bibitem{Burckhardt:2014POPL}
Burckhardt, S., Gotsman, A., Yang, H., Zawirski, M.:
Replicated data types: specification, verification, optimality.
\newblock In: Jagannathan, S., Sewell, P. (eds.) POPL 2014, pp. 271-284. ACM (2014)
\end{thebibliography}



\newpage

\appendix

\input{appendix}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
