%!TEX root = draft.tex
%\newcommand{\seqPQ}{\mathsf{SeqPQ}}

\section{Proving Distributed Linearizability}
\label{sec:proving distributed linearizability}

In this section, we propose the proving strategy for CRDT implementations based on simulation.



\subsection{Execution-Order Simulation Relation}
\label{subsec:time order of execution as linearization}


For some CRDT implementations, when proving its correctness, we can use the total-order of executions as linearization. For such CRDT implementations, our proof approach is as follows:


Given a object $\aobj$ and a specification \Spec{}, we construct a predicate $P(\mathit{config},\mathit{lin})$ where $\mathit{config} = (\gstates, \avisord, \downstreams)$ is a configuration of $\llbracket \aobj \rrbracket_{\mathit{op}}$, and $\mathit{lin}$ is a sequence used as linearization. We require $P(\mathit{config},\mathit{lin})$ to be a conjunction of the following statements:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Condition $C_1$ (linearizability): The history $(\alabelset, \avisord)$ is \crdtlinearizable{} w.r.t \Spec{} and $\mathit{lin}$ is a linearization. Here $\alabelset$ is the set of labels in domain of $\downstreams$.

\item[-] Condition $C_2$ (downstream): A statement about downstream for each operation label in $\mathit{config}$.

\item[-] Condition $C_3$ (sequential explanation): For each replica $\arep$, we have that $\gstates(\arep).\astate = \mathit{apply}(\mathit{lin},\gstates(\arep).\alabelset)$.
\end{itemize}

Here the function $\mathit{apply}(\mathit{lin},S)$ returns a local state obtained by applying downstream of operations in $S$ according to total order of $\mathit{lin}$.

Then, we need to prove that $P$ is a simulation relation. Moreover, we prove that we can always obtain a new linearization by putting a new operation after the tail of a old linearization. We call such relation a execution-order simulation relation.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $P(\mathit{config}_0,\epsilon)$ holds, where $\mathit{config}_0$ is the initial configuration.

\item[-] If $P(\mathit{config},\mathit{lin})$ holds and $\mathit{config} {\xrightarrow{\alabel}} \mathit{config}'$, then $P(\mathit{config}', \mathit{lin} \cdot \alabel)$ holds.

\item[-] If $P(\mathit{config},\mathit{lin})$ holds and $\mathit{config} {\xrightarrow{}} \mathit{config}'$, then $P(\mathit{config}',\mathit{lin})$ holds.
\end{itemize}

It is easy to see that, execution-order simulation relation implies a \crdtlinearizable{} relation proof with the execution as linearization.

We can see that, in the definition of $P(\mathit{config},\mathit{lin})$, condition $C_2$ is specific to CRDT implementations. The pseudo-code of or-set implementation is shown in \autoref{lst:or-set}. Its formal definition can be easily obtained from the pseudo-code and is omitted here.

\begin{lstlisting}[caption={Pseudo-code of the or-set CRDT}, captionpos=b,label={lst:or-set}]
  payload Set S, Timestamp maxTS
  initial S = @|$\emptyset$|@, maxTS = @|(0,myID())|@

  add(a) :
    atSource :
      precondition :
      assume maxTS = (c,r)
      let @|$\ats$|@ = (c+1,myID())
    downStream(a, @|$\ats$|@) :
      precondition:
      S = S @|$\cup$|@ {(a, @|$\ats$|@)}
      maxTS = max(maxTS, @|$\ats$|@)

  remove(a) :
    atSource :
      precondition : @|$\exists$|@ @|$\ats$|@, (a,@|$\ats$|@) @|$\in$|@ S
      let S@|$_1$|@ = @|$\{$|@ (a,@|$\ats$|@) @|$\in$|@ S @|$\}$|@
    downStream(S@|$_1$|@) :
      precondition :
      S = S @|$\setminus$|@ S@|$_1$|@

  read() :
    return {a @|$\vert$|@ @|$\exists$|@ ts, (a,@|$\ats$|@) @|$\in$|@ S}
\end{lstlisting}


In the downstream of a $\alabellongNoret[\mathit{add}]{\argv}$ operation, an tuple $(\argv,\ats)$ will be added to the local state; while in the downstream of a $\alabellongNoret[\mathit{add}]{\argv}$ operation, a set $S_1$ will be removed from the local state. We call such tuple $(\argv,\ats)$ or set $S_1$ the content of the downstream. The following is the condition $C_2$ for or-set implementation.

\begin{example}[Condition $C_2$ for or-set implementation]
\label{example:condition c2 for or-set implementation}
Given a configuration $(\gstates, \avisord, \downstreams)$ and an operation $\alabel \in \dom{\avisord}$, the downstream of $\alabel$ is given as follows: 

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\alabel = \alabellongNoret[\mathit{add}]{\argv}$: Assume $\alabel$ is generated for replica $\arep$. Then, the downstream of $\alabel$ is a function that maps a state $(S,\mathit{maxTS})$ into $(S \cup \{ (a,(c+1,r)) \}, \mathit{max}(\mathit{maxTS},(c+1,r)))$.  
    
    Here $c = \mathsf{max}\{ c' \vert \exists \alabel'$, such that $\alabel' = \alabellongNoret[\mathit{add}]{\_}, (\alabel', \alabel) \in \avisord$, and $(\_,(c',\_))$ is the downstream content of $\alabel' \}$. 

\item[-] If $\alabel = \alabellongNoret[\mathit{remove}]{\argv}$: The downstream of $\alabel$ is a function that maps a state $(S,\mathit{maxTS})$ into $(S \setminus S', \mathit{maxTS})$. 
    Here $S' = \{ \alabel' = \alabellongNoret[\mathit{add}]{\argv} \vert (\alabel',\alabel) \in \avisord, \neg \exists \alabel'' = \alabellongNoret[\mathit{remove}]{\argv}$, such that $(\alabel',\alabel''), (\alabel'',\alabel) \in \avisord \}$. 

\item[-] If $\alabel = \alabellongNoret[\mathit{read}]{}$: The downstream of $\downstreams(\alabel)$ is the identical function. 
\end{itemize}
\end{example}

The proof for or-set can be found in Appendix \ref{subsec:appendix proofs of or-set implementation}. The counter implementation can be similarly proved and its proof can be found in Appendix \ref{a}.




\subsection{Timestamp-Order Simulation Relation}
\label{subsec:time-stamp order as linearizabtion} 

For some CRDT implementation, such as RGA, the timestamp plays a important roles in conflict resolution. In such situation, the total-order of execution may not be used as linearization. For example, consider $\alabellongNoret[\mathit{write}]{c}$ and $\alabellongNoret[\mathit{add}]{d}$ actions of \figurename~\ref{fig:a failed example of composing a multi-value register with a last-write-win register} for LWW-register. Although $\alabellongNoret[\mathit{add}]{d}$ happens earlier in execution, since the timestamp of $\alabellongNoret[\mathit{write}]{c}$ is smaller than that of $\alabellongNoret[\mathit{write}]{d}$, we should put $\alabellongNoret[\mathit{write}]{c}$ before $\alabellongNoret[\mathit{write}]{d}$ in linearization.

Some method, such as $\mathit{write}$ of RGA, generates new timestamp; while other methods do not generate new timestamp. Based on this observation, we slightly change the operational semantics as follows: For the method that generate new timestamps, 


\[
  \inferrule[\text{\sc Operation-ts}]
  {\gstates(r) = (\alabelset, \astate) \\ \mathit{do}(\sigma,\amethod,\argv,\arep) = (\retv,\effector,\ats) \\  \effector(\astate) = \astate' \\ \alabel = \alabellongind{\argv}{\retv}{\ats} \\ \mathit{unique}(\ats) }
  {(\gstates, \avisord, \downstreams) \xrightarrow{\alabel} (\gstates[\arep \leftarrow (\alabelset \cup \{\alabel\}, \astate')],
    \avisord \cup (\alabelset \times \{\alabel\}), \downstreams[\alabel \rightarrow \effector])}
\]

Here we use $\mathit{do}(\sigma,\amethod,\argv,\arep) = (\retv,\effector,\ats)$ to emphasize that $\ats$ is the new-generated timestamp, and by $\alabel = \alabellongind{\argv}{\retv}{\ats}$ we record this timestamp in operation label. 

We require $P(\mathit{config},\mathit{lin})$ to be a conjunction of conditions $C_1$, $C_2$, $C_3$ and $C_4$, where condition $C_4$ is defined as follows: 

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Condition $C_4$ (timestamp): Given operations $\alabel$ and $\alabel'$ of methods that generate new timestamp. $(\alabel,\alabel') \in \avisord$ implies that the timestamp of $\alabel$ is less than that of $\alabel'$. 
\end{itemize} 

For simplicity, we define the notion of ``timestamp of operations'' as follows: 

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\alabel$ is an operation of methods that generate a new time-stamp $\ats$, then the timestamp of $\alabel$ is $\ats$. 

\item[-] If $\alabel$ is an operation of methods that do not generate new time-stamp: If $\alabel$ do not see any operation that generate new timestamp, then the timestamp of $\alabel$ is a default minimal timestamp; Else, the timestamp of $\alabel$ is $\mathsf{max} \{ \ats' \vert \exists \alabel'$, such that $(\alabel',\alabel) \in \avisord, \alabel$ generates a new time-stamp $\ats' \}$. 
\end{itemize} 

Then, we need to prove that $P$ is a simulation relation. Moreover, we prove that we can always obtain a new linearization by putting a new operation $\alabel$ after the last operation with a timestamp less or equal than $\alabel$. We call such relation a timestamp-order simulation relation. 

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $P(\mathit{config}_0,\epsilon)$ holds, where $\mathit{config}_0$ is the initial configuration.

\item[-] If $P(\mathit{config},\mathit{lin})$ holds and $\mathit{config} {\xrightarrow{\alabel}} \mathit{config}'$, then $P(\mathit{config}', \mathit{lin} \oplus \alabel)$ holds. Here  $\mathit{lin} \oplus \alabel$ is obtained from $\mathit{lin}$ by inserting $\alabel$ after the last operations with timestamp less or equal than that of $\alabel$. 

\item[-] If $P(\mathit{config},\mathit{lin})$ holds and $\mathit{config} {\xrightarrow{}} \mathit{config}'$, then $P(\mathit{config}',\mathit{lin})$ holds.
\end{itemize}

It is easy to see that, timestamp-order simulation relation implies a \crdtlinearizable{} relation proof. Moreover, we can obtain a linearization in the following form: For each operation $\alabel$, the timestamp of $\alabel$ is larger or equal than that of operations before $\alabel$ in linearization, and is less or equal than that of operations after $\alabel$ in linearization. We call such linearization a timestamp linearization. We can see that timestamp-order simulation relation implies a timestamp linearization.

The following is the condition $C_2$ for RGA. 

\begin{example}[Condition $C_2$ for RGA]
\label{example:condition c2 for rga} 
Given a configuration $(\gstates, \avisord, \downstreams)$ and an operation $\alabel \in \dom{\avisord}$, the downstream of $\alabel$ is given as follows. 

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\alabel = \alabellongNoret[\mathit{add}]{a,b}$: Assume $\alabel$ is generated for replica $\arep$. Then, the downstream of $\alabel$ is a function that maps a state $(N,\mathit{Tomb})$ into $(N \cup \{ (a,\ats_b,\ats_a) \}, \mathit{Tomb})$. 
    
    If $a = \circ$, then $\ats_a = (0,r_0)$; Else, $\ats_a$ is the timestamp of $\alabellongNoret[\mathit{add}]{\_,a}$. If no $\mathit{add}$ operation is visible to $\alabel$, then $\ats_b = (1,\arep)$; Else, $\ats_b = (c+1,\arep)$, and $c = \mathsf{max}\{ c' \vert \exists \alabel' = \alabellongNoret[\mathit{add}]{\_,\_}$, such that $(\alabel', \alabel) \in \avisord$, and $(\_,(c',\_))$ is the timestamp of $\alabel' \}$. 


\item[-] If $\alabel = \alabellongNoret[\mathit{remove}]{\argv}$: Then, the downstream of $\alabel$ is a function that maps a state $(N,\mathit{Tomb})$ into $(N, \mathit{Tomb} \cup \{ a \})$.

\item[-] If $\alabel = \alabellongNoret[\mathit{read}]{}$: Then, $\downstreams(\alabel)$ is the identical function. 
\end{itemize} 
\end{example}

The detailed of RGA can be found in Appendix \ref{subsec:appendix proofs of rga}. The LWW register can be similarly proved and its proof can be found in Appendix \ref{a}.











\forget{
\section{Proving Distributed Linearizability}
\label{sec:proving distributed linearizability}

In this section, we propose the proving strategy for CRDT implementations based on simulation.



\subsection{Time Order of Executions as Linearization}
\label{subsec:time order of execution as linearization}


For some CRDT implementations, the time-order of operations in execution is a linearization. For such CRDT implementation, our proof approach is as follows:

Given a object $\mathit{obj}$ and a specification $\mathit{spec}$, we construct a predicate $P(\mathit{config},h,\mathit{lin},\mathit{map})$ where $\mathit{config} = (R,T,\mathit{MsgHB},\mathit{MsgDel})$ is a configuration of $\llbracket \mathit{obj} \rrbracket_{\mathit{op}}$, $h$ is a history, $\mathit{lin} \in \mathbb{A}^*$ is sequence used as linearization, and $\mathit{map} \subseteq \mathbb{MID} \times \mathbb{OID}$ is a function used to associate messages of $\mathit{config}.T$ with operations of $h$. We require $P(\mathit{config},h,\mathit{lin},\mathit{map})$ to be a conjunction of the following statements:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $C_1$ (linearizability): $h$ is distributed linearizable w.r.t $\mathit{spec}$ and $\mathit{lin}$ is a linearization.

\item[-] $C_2$ (correct $\mathit{map}$): $\mathit{map}$ is a bijection between messages of $\mathit{config}.T$ and operations of $h$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] For message visibility of $\mathit{config}$: $(o_1,o_2) \in h.\mathit{vis}$, if and only if $(\mathit{map}(o_1),\mathit{map}(o_2)) \in \mathit{config}.\mathit{MsgHB}$.

    \item[-] For message delivery of $\mathit{config}$: If $(\mathit{mid},r) \notin \mathit{config}.\mathit{MsgDel}$, then for each $\mathit{mid}'$ where the message of $\mathit{mid}'$ is of replica $r$, $(\mathit{map}(\mathit{mid}),\mathit{map}(\mathit{mid}')) \notin h.\mathit{vis}$.
    \end{itemize}

\item[-] $C_3$ (causal delivery): $h.\mathit{vis}$ is transitive. Moreover, if $(o_1,o_2) \in h.\mathit{vis}$ and $(\mathit{map}(o_2),r) \in \mathit{config}.\mathit{MsgDel}$, then $(\mathit{map}(o_1),r) \in \mathit{config}.\mathit{MsgDel}$.

\item[-] $C_4$ (message content): A statement about message content of each message in $\mathit{config}.T$.

\item[-] $C_5$ (sequential explanation): For each replica $r$, we have that $\mathit{config}.R(r) = \mathit{apply}(\mathit{lin},\mathit{vd}(h,\mathit{config},r))$.
\end{itemize}

Here $\mathit{vd}(h,\mathit{config},r) = \{ o \vert (o,o') \in h.\mathit{vis}, \mathit{map}(o')$ is of replica $r \} \cup \{ o \vert (\mathit{map}(o),r) \in \mathit{config}.\mathit{MsgHB} \}$. $\mathit{vd}(h,\mathit{config},r)$ is the set of operations that are either visible to some operation of replica $r$, or has been delivered into replica $r$. The function $\mathit{apply}(\mathit{lin},S)$ returns a local state by applying messages of operations in $S$ according to total order $\mathit{lin}$.


Then, we need to prove that $P$ is a simulation relation in a way shown below. Note that here we choose the time order of executions as the linearization. Or we can say, we obtain new linearization by inserting a new operation after the tail of the old linearization.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $P(\mathit{config}_0,\epsilon,\epsilon,\emptyset)$ holds, where $\mathit{config}_0$ is the initial configuration.

\item[-] If $P(\mathit{config},h,\mathit{lin},\mathit{map})$ holds and $\mathit{config} {\xrightarrow{\mathit{do}(m,a,b,r,\mathit{mid})}} \mathit{config}'$, then $P(\mathit{config}', h' = h \otimes (m(a) \Rightarrow b,i,\mathit{obj}), \mathit{lin} \cdot (m(a) \Rightarrow b,i,h'.\mathit{vis}^{-1}(i)),\mathit{map} \cup \{ (\mathit{mid}, i) \})$ holds. Here $i$ is a unique operation identifier.

\item[-] If $P(\mathit{config},h,\mathit{lin},\mathit{map})$ holds and $\mathit{config} {\xrightarrow{\mathit{do}(m,a,b,r)}} \mathit{config}'$, then $P(\mathit{config}',h' = h \otimes (m(a) \Rightarrow b,i,\mathit{obj}), \mathit{lin} \cdot (m(a) \Rightarrow b,i,h'.\mathit{vis}^{-1}(i)),\mathit{map})$ holds. Here $i$ is a unique operation identifier.

\item[-] If $P(\mathit{config},h,\mathit{lin},\mathit{map})$ holds and $\mathit{config} {\xrightarrow{\mathit{receive}(\mathit{mid},r)}} \mathit{config}'$, then $P(\mathit{config}',h,\mathit{lin},\mathit{map})$ holds.
\end{itemize}

Given history $h = (\mathit{Op},\mathit{ro},\mathit{vis})$ and operation $o = (m(a) \Rightarrow b,i,\mathit{obj})$, $h \otimes o$ returns a history $(\mathit{Op}',\mathit{ro}',\mathit{vis}')$, where $\mathit{Op}' = \mathit{Op} \cup \{ o \}$, $\mathit{ro}' = \mathit{ro} \cup \{ (o',o) \vert \mathit{map}(o')$ is a message of replica $r \}$, and $\mathit{vis}' = (\mathit{vis} \cup \{ (o',o) \vert (\mathit{map}(o'),r) \in \mathit{config}.\mathit{MsgDel} \} \cup \{ (o',o) \vert \mathit{map}(o')$ is a message of replica $r \})^*$. A predicate $P(\mathit{config},h,\mathit{lin},\mathit{map})$ is called an invariant, if it satisfies the simulation relation described above. It is obvious that the existence of an invariant implies distributed linearizability.

%The following lemma states that the existence of such invariant implies distributed linearizability, and each sequence consistent with visibility is a linearization. To prove this lemma, we first prove that by induction on length of executions, we obtain a linearization. Then by definition of t0-specification, we know that each sequence consistent with visibility is a linearization.

%\begin{lemma}
%\label{lemma:invariant of operation-based CRDT implies distributed linearizability for t0-specification}
%If there exists an invariant $P$ for a CRDT object $\mathit{obj}$ and a t0-specification $\mathit{spec}$, then each history of $\mathit{history}(\llbracket \mathit{obj} \rrbracket_{\mathit{op}})$ is distributed linearizable w.r.t $\mathit{spec}$, and each sequences that consistent with visibility is a linearization of $h$.
%\end{lemma}

In definition of invariant, only condition $C_4$ is specific to CRDT implementation. The following is the condition $C_4$ for or-set implementation.

\begin{example}[$C_4$ for or-set implementation]
\label{example:c4 for or-set implementation}

Given history $h = (\mathit{Op},\mathit{ro},\mathit{vis})$ and a update operation $o$ of $h$, the message of $o$ is given as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $o$ is a $\mathit{add}(a)$ operation of replica $r$:  then the message of $o$ is $(a,(c+1,r))$, where $c = \mathit{max}\{ c_1 \vert \exists o' = (\_,\_,\_,\mathit{ts}), (o',o) \in \mathit{vis}, \mathit{ts} = (c_1,\_) \}$.

\item[-] If $o$ is a $\mathit{rem}(a)$ operation: the message of $o$ is $S_1$, where $S_1 = \{ (a,\mathit{ts}') \vert (a,\mathit{ts}') \in S(o) \}$. We also require that $S_1 \neq \emptyset$.
\end{itemize}
Here $S(o) = \{ (b,\mathit{ts}') \vert b \in D, \exists o' = (\mathit{add}(b),\_,\_,\mathit{ts}')$, $(o',o) \in \mathit{vis}$, for each $\mathit{rem}(b)$ operation $o'' \neq o, (o'',o) \in \mathit{vis} \Rightarrow (o',o'') \neq \mathit{vis} \}$.
\end{example}

The proof for or-set can be found in Appendix \ref{subsec:appendix proofs of or-set implementation}. The counter implementation can be similarly proved and its proof can be found in Appendix \ref{a}.




\subsection{Time-Stamp Order as Linearization}
\label{subsec:time-stamp order as linearizabtion}

For some CRDT implementation, such as RGA, time-stamp is used for conflict resolution, and it can not be only use the time order of executions as linearization. For example, there must be a fixed order for concurrent $\mathit{add}(b,a)$ and $\mathit{add}(c,a)$, and this order may contradict the time order of execution.

For such CRDT implementation, we can assume that each operation also gives the information of time-stamp. Or we can say, we can implicitly assume that each operation is of the form $o = (\ell,i,\mathit{obj},\mathit{ts})$, where $\mathit{ts}$ is the ``time-stamp'' of this operation:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $o$ generates a new unique time-stamp, then $\mathit{ts}$ is this new time-stamp.

\item[-] If $o$ does not generate new time-stamp, then $\mathit{ts}$ is the maximum of time-stamp among operations visible to $o$.

\item[-] Moreover, we require that given operations $o_1$ and $o_2$, if $(o_1,o_2) \in \mathit{vis}$, then the time-stamp of $o_1$ is less or equal than that of $o_2$.
\end{itemize}

Then, for such CRDT implementation, we use the time-stamp order as linearization and prove distributed linarization as follows:

We use a same predicate $P(\mathit{config},h,\mathit{lin},\mathit{map})$ as in subsection \ref{subsec:proof approach for t0-specifications}. We still need to prove that $P$ is a simulation relation. The difference is the process of constructing new linearization: a operation with time-stamp $\mathit{ts}$ is put after the last operation with a time-stamp less or equal than $\mathit{ts}$ in old linearization.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $P(\mathit{config}_0,\epsilon,\epsilon,\emptyset)$ holds, where $\mathit{config}_0$ is the initial configuration.

\item[-] If $P(\mathit{config},h,\mathit{lin},\mathit{map})$ holds and $\mathit{config} {\xrightarrow{\mathit{do}(m,a,b,r,\mathit{mid})}} \mathit{config}'$, then $P(\mathit{config}', h' = h \otimes (m(a) \Rightarrow b,i,\mathit{obj}),\mathit{lin}',\mathit{map} \cup \{ (\mathit{mid}, i) \})$ holds. Here $i$ is a unique operation identifier, and $\mathit{lin}'$ is obtained from $\mathit{lin}$ by inserting $(m(a) \Rightarrow b,i,h'.\mathit{vis}^{-1}(i))$ after the last operation with time-stamp less or equal than $\mathit{ts}$.

\item[-] If $P(\mathit{config},h,\mathit{lin},\mathit{map})$ holds and $\mathit{config} {\xrightarrow{\mathit{do}(m,a,b,r)}} \mathit{config}'$, then $P(\mathit{config}',h' = h \otimes (m(a) \Rightarrow b,i,\mathit{obj}),\mathit{lin}',\mathit{map})$ hold. Here $i$ is a unique operation identifier, and $\mathit{lin}'$ is obtained from $\mathit{lin}$ by inserting $(m(a) \Rightarrow b,i,h'.\mathit{vis}^{-1}(i))$ after the last operation with time-stamp less or equal than $\mathit{ts}$.

\item[-] If $P(\mathit{config},h,\mathit{lin},\mathit{map})$ holds and $\mathit{config} {\xrightarrow{\mathit{receive}(\mathit{mid},r)}} \mathit{config}'$, then $P(\mathit{config}',h,\mathit{lin},\mathit{map})$ holds.
\end{itemize}

A predicate $P(\mathit{config},h,\mathit{lin},\mathit{map})$ is called an invariant, if it satisfies the simulation relation described above. It is obvious that the existence of an invariant implies distributed linearizability.

%The following lemma states that the existence of such invariant implies distributed linearizability, and each sequence consistent with visibility and time-stamp is a linearization. To prove this lemma, we first prove that by induction on length of executions, we obtain a linearization. Then by definition of t1-specification, we know that each sequence consistent with visibility and time-stamp is a linearization.

%\begin{lemma}
%\label{lemma:invariant of operation-based CRDT implies distributed linearizability for t1-specification}
%If there exists an invariant $P$ for a CRDT object $\mathit{obj}$ and a t1-specification $\mathit{spec}$, then each history of $\mathit{history}(\llbracket \mathit{obj} \rrbracket_{\mathit{op}})$ is distributed linearizable w.r.t $\mathit{spec}$, and each sequences that consistent with visibility and time-stamp is a linearization of $h$.
%\end{lemma}


In definition of invariant, only condition $C_4$ is specific to CRDT implementation. The following is the condition $C_4$ for RGA.

\begin{example}[$C_4$ for RGA]
\label{example:c4 for rga}
Given history $h = (\mathit{Op},\mathit{ro},\mathit{vis})$ and a update operation $o$ of $h$, the message of $o$ is given as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $o$ is a $\mathit{add}(a,b)$ operation of replica $r$: the message of $o$ is $(a,\mathit{ts}_a,\mathit{ts}_b)$, where $\mathit{ts}_b$ is the time-stamp of operation $\mathit{add}(b,\_)$ in $h$, $\mathit{ts}_a = (c+1, r)$, and $c = \mathit{max}\{ c_1 \vert \exists o' = (\_,\_,\_,\mathit{ts}), (o',o) \in \mathit{vis}, \mathit{ts} = (c_1,\_) \}$.

\item[-] If $o$ is a $\mathit{rem}(a)$ operation: the message of $o$ is $a$.
\end{itemize}
\end{example}

The detailed of RGA can be found in Appendix \ref{subsec:appendix proofs of rga}. The LWW register can be similarly proved and its proof can be found in Appendix \ref{a}.
}







%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft"
%%% End:
